The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-11-19 script/en/chapters/digitalsignatures.tex
CT script: Minor changes (more consistent, more up-to-date)

\newpage
% --------------------------------------------------------------------------
\hypertarget{Chapter_Hashes-and-Digital-Signatures}{}
\chapter{Hash Functions and Digital Signatures}
\label{Chapter_Hashes-and-Digital-Signatures}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Koy H., June 2002;
Updates: Feb. 2003, June 2005, July 2009)
% --------------------------------------------------------------------------
%\vskip + 30pt
\begin{ctsquote}
We can make everything out of this world, but we cannot create a world,
where humans in some ten thousand years can think: 'Ok, now it is enough.
Everything should stay like it is. Let's do no changes any more, don't do
inventions any more, because it cannot become better, and if, then we
don't want this.'
\caption[Stanislaw Lem]{Stanislaw Lem\footnotemark}\index{Lem, Stanislaw}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{This was the answer of Stanislaw Lem
to heavy critics at his philosophical main book ~``Summa Technologiae'',
1964, where he thought about the possibility of an evolution creating
artificial intelligence.\\}

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} you can also generate and check 
digital signatures: Using\\
the submenus of the main menu {\bf Digital Signatures / PKI}  or using\\
menu {\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem 
\textbackslash{} Signature Demonstration (Signature Generation)}.
}.

In detail, a \index{Signature procedure} {\em Signature procedure} looks like
this: \\ Senders use their message and secret key to calculate the digital
signature for the message. Compared to hand-written signatures, digital
signatures therefore have the advantage that they also depend on the document to
be signed. Signatures from one and the same participant are different unless the
signed documents are completely identical. Even inserting a blank in the text
would lead to a different signature. The recipient of the message would
therefore detect any injury to the message integrity as this would mean that the
signature no longer matches the document and is shown to be incorrect when
verified.

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. The procedure we just described
has in practice, however, a decisive disadvantage. The signature would be
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we apply
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CrypTool\index{CrypTool} at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Analysis \textbackslash{} Hash}
you have the possibilities
% hier die items nicht einruecken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value,\\
\item to calculate a key from a password according to the PKCS\#5 
      standard\index{PKCS\#5}, \\
\item to calculate HMACs from a text and a secret key, and\\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
} to the document -- before signing. The output of the hash function will 
then be signed.

@@

  With CrypTool\index{CrypTool} you can comprehend the birthday attack
  \index{Attack!birthday} on digital signatures: \\
  using the menu {\bf Analysis \textbackslash{} Hash \textbackslash{} 
  Attack on the Hash Value of the Digital Signature}.
  }  
or \index{RIPEMD-160} RIPEMD-160\footnote{%
  RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
  identifiers defined by the ISO-identified organization TeleTrusT, both 
  as hash algorithm and in combination with RSA.
  RIPEMD-160 is also part of the ISO/IEC international standard 
  ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
  RIPEMD-128 and SHA-1.
  Further details: \\
- \href{http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html}
   {\tt http://www.esat.kuleuven.ac.be/\~{}bosselae/ripemd160.html}\\
- \href{http://www.ietf.org/rfc/rfc2857.txt}
   {\tt http://www.ietf.org/rfc/rfc2857.txt} (``The Use of HMAC-RIPEMD-160-96
   within ESP and AH'').
  }%
) were trusted.

At Crypto 2004 (August 2004)\footnote{%
    \href{http://www.iacr.org/conferences/crypto2004/}
 {\texttt{http://www.iacr.org/conferences/crypto2004/}} }
this safety-feeling was disputed: 
Chinese researchers published collision attacks against MD4, SHA-0 and
parts of SHA-1. This globally caused new motivation to engage in
new hash attack methods.

@@

Already before Crypto 2004 the U.S. National Institute of Standards and
Technology (NIST)\index{NIST} announced, to discontinue SHA-1 in the next few
years. So it is recommended not to use SHA-1 for new products generating digital
signatures. The SHA-2 family \cite{ds:FIPS180-3} provides stronger algorithms. To
address new findings in cryptanalysis, NIST has opened a competition to develop a
new cryptographic hash algorithm ``SHA-3'' in 2008. The competition is expected
to end in 2012.\footnote{\url{http://csrc.nist.gov/groups/ST/hash/sha-3/}}

Further information about this topic can be found in the article
``Hash cracked -- The consequences of the successful attacks on SHA-1''
by Reinhard Wobst and J\"urgen Schmidt\footnote{% 
      \href{http://www.heise.de/security/artikel/56634}
   {\texttt{http://www.heise.de/security/artikel/56634}}. \\
  Further references are e.g.: \\
      \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
   {\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}} \\
      \href{http://csrc.nist.gov/CryptoToolkit/tkhash.html}
   {\texttt{http://csrc.nist.gov/CryptoToolkit/tkhash.html}}.
} 
  by Heise Security.

@@

The signature procedure with hash functions is as follows:\footnote{% 
Compare: \href{http://en.wikipedia.org/wiki/Digital\_signature}
      {\texttt{http://en.wikipedia.org/wiki/Digital\_signature}}.
} \\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

\bibitem[FIPS186-2]{ds:FIPS186-2} U.S. Department of Commerce/N.I.S.T. ,
    \index{FIPS186-2} \\
    {\em Digital Signature Standard (DSS)}, \\
    January 27, 2000. Change Note: October 5, 2001.\\
  \href{http://csrc.nist.gov/publications/fips/fips186-2/fips186-2-change1.pdf}
   {\tt http://csrc.nist.gov/publications/fips/fips186-2/fips186-2-change1.pdf}

>>>>

\newpage
% --------------------------------------------------------------------------
\hypertarget{Chapter_Hashes-and-Digital-Signatures}{}
\chapter{Hash Functions and Digital Signatures}
\label{Chapter_Hashes-and-Digital-Signatures}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Koy H., June 2002;
Updates: Feb. 2003, June 2005, July 2009, Nov. 2012)
% --------------------------------------------------------------------------
%\vskip + 30pt
\begin{ctsquote}
We can make everything out of this world, but we cannot create a world,
where humans in some ten thousand years can think: 'Ok, now it is enough.
Everything should stay like it is. Let's do no changes any more, don't do
inventions any more, because it cannot become better, and if, then we
don't want this.'
\caption[Stanislaw Lem]{Stanislaw Lem\footnotemark}\index{Lem, Stanislaw}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{This was the answer of Stanislaw Lem
to heavy critics at his philosophical main book ~``Summa Technologiae'',
1964, where he thought about the possibility of an evolution creating
artificial intelligence.}

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key.

In detail, a \index{Signature procedure} {\em Signature procedure}\footnote{%
With CT1\index{CrypTool 1} you can also generate and check 
digital signatures: Using\\
the submenus of the main menu {\bf Digital Signatures / PKI}  or using\\
menu {\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem 
\textbackslash{} Signature Demonstration (Signature Generation)}.\\
Also with JCT\index{JCrypTool} (in the default and the functional perspektive)
its possible to create different kinds of electronic signatures.
} looks like
this: \\ Senders use their message and secret key to calculate the digital
signature for the message. Compared to hand-written signatures, digital
signatures therefore have the advantage that they also depend on the document to
be signed. Signatures from one and the same participant are different unless the
signed documents are completely identical. Even inserting a blank in the text
would lead to a different signature. The recipient of the message would
therefore detect any injury to the message integrity as this would mean that the
signature no longer matches the document and is shown to be incorrect when
verified.

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. The procedure we just described
has in practice, however, a decisive disadvantage. The signature would be
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we apply
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CT1\index{CrypTool 1} at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Analysis \textbackslash{} Hash}
you have the possibilities
% hier die items nicht einruecken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value,\\
\item to calculate a key from a password according to the PKCS\#5 
      standard\index{PKCS\#5}, \\
\item to calculate HMACs from a text and a secret key, and\\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
CT2\index{CrypTool 2} and JCT\index{JCrypTool} also contain different hash
methods. }
to the document -- before signing. The output of the hash function will 
then be signed.

@@

  With CT1\index{CrypTool 1} you can comprehend the birthday attack
  \index{Attack!birthday} on digital signatures: \\
  using the menu {\bf Analysis \textbackslash{} Hash \textbackslash{} 
  Attack on the Hash Value of the Digital Signature}.\\
  CT2\index{CrypTool 2} contains an MD5 collider.
  }  
or \index{RIPEMD-160} RIPEMD-160\footnote{%
  RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
  identifiers defined by the ISO-identified organization TeleTrusT, both 
  as hash algorithm and in combination with RSA.
  RIPEMD-160 is also part of the ISO/IEC international standard 
  ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
  RIPEMD-128 and SHA-1.
  Further details: \\
- \url{http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html}\\
- \url{http://www.ietf.org/rfc/rfc2857.txt} (``The Use of HMAC-RIPEMD-160-96
   within ESP and AH'').
  }%
) were trusted.

At Crypto 2004 (August 2004)\footnote{%
    \url{http://www.iacr.org/conferences/crypto2004/} }
this safety-feeling was disputed: 
Chinese researchers published collision attacks against MD4, SHA-0 and
parts of SHA-1. This globally caused new motivation to engage in
new hash attack methods.

@@

Already before Crypto 2004 the U.S. National Institute of Standards and
Technology (NIST)\index{NIST} announced, to discontinue SHA-1 in the next few
years. So it is recommended not to use SHA-1 for new products generating digital
signatures. The SHA-2 family \cite{ds:FIPS180-3} provides stronger algorithms.
To address new findings in cryptanalysis, NIST has opened a competition to develop a new cryptographic hash algorithm in 2008: In October 2012
Keccak\index{Keccak} was announced as ``SHA-3''.\footnote{%
\url{http://csrc.nist.gov/groups/ST/hash/sha-3/}}

Further information about this topic can be found in the article
``Hash cracked -- The consequences of the successful attacks on SHA-1''
by Reinhard Wobst and J\"urgen Schmidt\footnote{% 
      \url{http://www.heise.de/security/artikel/56634}. \\
  Further references are e.g.: \\
      \url{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}\\
      \url{http://csrc.nist.gov/CryptoToolkit/tkhash.html}.
} 
  by Heise Security.

@@

The signature procedure with hash functions\footnote{% 
Compare: \url{http://en.wikipedia.org/wiki/Digital\_signature}.
} is as follows: Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

\bibitem[FIPS186-2]{ds:FIPS186-2} U.S. Department of Commerce/N.I.S.T. ,
    \index{FIPS186-2} \\
    {\em Digital Signature Standard (DSS)}, \\
    January 27, 2000. Change Note: October 5, 2001.\\
   \url{http://csrc.nist.gov/publications/fips/fips186-2/fips186-2-change1.pdf}

### 2012-11-19 script/en/chapters/ellipticcurves.tex
CT script: Minor changes (more consistent, more up-to-date)

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths. If we assume that the computing power increases
by Moore's law (i.~e.\ it doubles every 18 months)\footnote{empirical
knowledge by Gordon Moore, co-founder of Intel, 1965}, then the evolution of
the key lengths for secure communication will be as figure~\ref{RSAKeylength}
\cite{ec:Lenstra1999} (source: Arjen Lenstra and Eric Verheul:
\href{http://cryptosavvy.com/table.htm}
{\texttt{http://cryptosavvy.com/table.htm}}).

@@

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field
has characteristic $2$.)\footnote{An animation of the addition of points
on elliptic curves can be found on the Certicom\index{Certicom} homepage \\
\url{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}}

@@

There are factorization%
\footnote{Especially John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorization algorithms; also at
factorization with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics'':
\href{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}
{\texttt{http://www.eff.org/Privacy/Crypto\_misc/DESCracker/HTML/19990118\_rsa\_awards.html}}.
}
algorithms based on elliptic curves%
\footnote{In 1987 H.W. Lenstra published
a factorization algorithm, based on elliptic curves (see \cite{ec:Lenstra1987}).
The biggest compound number currently 
factorized\index{Factorization!factoring records} with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

% -----------------------------------------------------------------------------
\subsection{CrypTool}

CrypTool offers elliptic curves for the digital signature function%
\footnote{%
The dialog box, which appears in CrypTool\index{CrypTool} after clicking the menu
{\bf Digital Signatures/PKI \textbackslash{} Sign Message},
offers the EC methods ECSP-DSA and ECSP-NR.

} and for ECC-AES hybrid encryption%
\footnote{%
Within CrypTool\index{CrypTool} you can find this technique using the menu
path {\bf Crypt \textbackslash{} Hybrid}.

@@

@@

    \bibitem[Cassels1991]{ec:Cassels1991} J. W. S. Cassels,
	\index{Cassels 1991} \\
        {\em Lectures on elliptic curves},\\
	Cambridge University Press, 1991, 143 pages.
    \bibitem[Koblitz1984]{ec:Koblitz1984}N. Koblitz, 
	\index{Koblitz 1984} \\
        {\em Introduction to elliptic curves and modular forms},\\
	Graduate Texts in Mathematics, Springer-Verlag, 1984.
    \bibitem[Koblitz1998]{ec:Koblitz1998}N. Koblitz,  
	\index{Koblitz 1998}\\
	{\em Algebraic aspects of Cryptography. With an appendix on 
	Hyperelleptic curves by Alfred J. Menezes, Yi Hong Wu and Robert 
	J. Zuccherato}, \\
	Springer-Verlag, 1998, 206 pages.
    \bibitem[Lenstra1987]{ec:Lenstra1987} H.W. Lenstra, 
        \index{Lenstra 1987} \\
        {\em Factoring integers with elliptic curves}, \\
	Annals of Mathematics 126, pp. 649-673, 1987.
    \bibitem[Lenstra1999]{ec:Lenstra1999} Arjen K. Lenstra, Eric R. Verheul
        \index{Lenstra/Verheul 1999} \\
        {\em Selecting Cryptographic Key Sizes (1999)},\\
	Journal of Cryptology: the journal of the International 
	Association for Cryptologic Research \\
	\href{http://www.cryptosavvy.com/cryptosizes.pdf}
	{\texttt{http://www.cryptosavvy.com/cryptosizes.pdf}}
    \bibitem[Menezes1993]{ec:Menezes1993}A. J. Menezes, 
	\index{Menezes 1993} \\
        {\em Elliptic curve public key cryptosystems},\\
	Kluwer Academic Publishers, 1993.
    \bibitem[Silverman1986]{ec:Silverman1986} J. Silverman, 
        \index{Silverman 1986}\\
        {\em The Arithmetic of Elliptic Curves},\\
	Springer-Verlag, 1986.
    \bibitem[Silverman1992]{ec:Silverman1992}J. Silverman,
	\index{Silverman 1992} \\
        {\em The arithmetic of elliptc curves},\\
	Graduate Texts in Mathematics, Springer-Verlag, 1992.
    \bibitem[SilvermanTate1992]{ec:SilvermanTate1992}J. Silverman, J. Tate,
	\index{Silverman/Tate 1992} \\
        {\em Rational points on elliptic curves},\\
	Springer-Verlag, 1992.

@@

   \item Certicom Online Tutorial\index{Certicom},\\
        \url{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}

>>>>

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths. If we assume that the computing power increases
by Moore's law (i.~e.\ it doubles every 18 months)\footnote{empirical
knowledge by Gordon Moore, co-founder of Intel, 1965}, then the evolution of
the key lengths for secure communication will be as figure~\ref{RSAKeylength}
\cite{ec:Lenstra1999} (source: Arjen Lenstra and Eric Verheul:
\url{http://cryptosavvy.com/table.htm}).

@@

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field
has characteristic $2$.)\footnote{%
An animation of the addition of points
on elliptic curves can be found on the Certicom\index{Certicom} homepage \\
\url{http://www.certicom.com/index.php/ecc-tutorial}\\
See also the web link about the \hyperlink{ec:Web-Link:Java_Laubrock}
{Java-Tutorial} at the end of this chapter.
}

@@

There are factorization%
\footnote{Especially John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorization algorithms; also at
factorization with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics'':
\url{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}.\\
In 1987 H.W. Lenstra published an often used factorization algorithm,
based on elliptic curves (see \cite{ec:Lenstra1987}).
}
algorithms based on elliptic curves%
\footnote{The biggest compound numbers currently 
factorized\index{Factorization!factoring records} with elliptic curves have
about 80 decimal digits:\\
\url{http://www.loria.fr/~zimmerma/records/top50.html}.\\
See also the web link about the \hyperlink{Lenstra2}{ECMNET project}\index{ECMNET} at the end of this chapter.
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

% -----------------------------------------------------------------------------
\subsection{CrypTool}\index{CrypTool}

CT1 offers elliptic curves for the digital signature function%
\footnote{%
The dialog box, which appears in CT1\index{CrypTool 1} after clicking the menu
{\bf Digital Signatures/PKI \textbackslash{} Sign Message},
offers the EC methods ECSP-DSA and ECSP-NR.

} and for ECC-AES hybrid encryption%
\footnote{%
Within CT1\index{CrypTool 1} you can find this technique using the menu
path {\bf Crypt \textbackslash{} Hybrid}.

@@

Step-by-step point addition on elliptic curves is visualized in CT1\index{CrypTool 1} and JCT\index{JCrypTool}.%
\footnote{%
CT1: menu {\bf Digital Signatures/PKI \textbackslash{} Signature Demonstration (Signature Generation)},\\
JCT (default perspective): menu {\bf Visuals \textbackslash{} Elliptic Curve Calculations}.

}

@@

    \bibitem[Cassels1991]{ec:Cassels1991} J. W. S. Cassels,
	\index{Cassels 1991} \\
        {\em Lectures on elliptic curves},\\
	Cambridge University Press, 1991, 143 pages.
    \bibitem[Koblitz1984]{ec:Koblitz1984}N. Koblitz, 
	\index{Koblitz 1984} \\
        {\em Introduction to elliptic curves and modular forms},\\
	Graduate Texts in Mathematics, Springer-Verlag, 1984.
    \bibitem[Koblitz1998]{ec:Koblitz1998}N. Koblitz,  
	\index{Koblitz 1998}\\
	{\em Algebraic aspects of Cryptography. With an appendix on 
	Hyperelleptic curves by Alfred J. Menezes, Yi Hong Wu and Robert 
	J. Zuccherato}, \\
	Springer-Verlag, 1998, 206 pages.
    \bibitem[Lenstra1987]{ec:Lenstra1987} H.W. Lenstra, 
        \index{Lenstra 1987} \\
        {\em Factoring integers with elliptic curves}, \\
	Annals of Mathematics 126, pp. 649-673, 1987.
    \bibitem[Lenstra1999]{ec:Lenstra1999} Arjen K. Lenstra, Eric R. Verheul
        \index{Lenstra/Verheul 1999} \\
        {\em Selecting Cryptographic Key Sizes (1999)},\\
	Journal of Cryptology: the journal of the International 
	Association for Cryptologic Research \\
	\url{http://www.cryptosavvy.com/cryptosizes.pdf}
    \bibitem[Menezes1993]{ec:Menezes1993}A. J. Menezes, 
	\index{Menezes 1993} \\
        {\em Elliptic curve public key cryptosystems},\\
	Kluwer Academic Publishers, 1993.
    \bibitem[Silverman1986]{ec:Silverman1986} J. Silverman, 
        \index{Silverman 1986}\\
        {\em The Arithmetic of Elliptic Curves},\\
	Springer-Verlag, 1986.
    \bibitem[Silverman1992]{ec:Silverman1992}J. Silverman,
	\index{Silverman 1992} \\
        {\em The arithmetic of elliptc curves},\\
	Graduate Texts in Mathematics, Springer-Verlag, 1992.
    \bibitem[SilvermanTate1992]{ec:SilvermanTate1992}J. Silverman, J. Tate,
	\index{Silverman/Tate 1992} \\
        {\em Rational points on elliptic curves},\\
	Springer-Verlag, 1992.

@@

   \hypertarget{ec:Web-Link:Certicom_Tutorial}{} 
   \item Certicom Online Tutorial\index{Certicom},\\
        \url{http://www.certicom.com/index.php/ecc-tutorial}

   \hypertarget{ec:Web-Link:Java_Laubrock}{} 
   \item Tutorial with Java applets -- Crypto methods based on elliptic curves,
        thesis by Thomas Laubrock, 1999 (German only),\\
        \url{http://www.warendorf-freckenhorst.de/elliptische-kurven/frame.html}

### 2012-11-19 script/en/chapters/menus.tex
CT script: Minor changes (more consistent, more up-to-date)

\noindent Within the Startcenter you have three main ways to use the implemented functionality:
\begin{itemize}
   \item the Wizard, which leads you to the functions.
   \item the Workspace, where you can compose the components by yourself according to the visual programming\index{Visual programming} concept.
   \item the template tree, which offers ready-to-run workflows.
 \end{itemize}

>>>>

\noindent Within the Startcenter you have three main ways to use the implemented functionality:
\begin{itemize}
   \item via the Wizard, which leads you to the functions.
   \item via the Workspace, where you can compose the components (e.g. an encryption method, a text input function, ...)  by yourself according to the visual programming\index{Visual programming} concept.
   \item via the template tree, which offers ready-to-run workflows.
 \end{itemize}
